Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-02-04 paper/rebuttal/rebuttal.tex

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\DeclareUnicodeCharacter{FB01}{fi}
\usepackage{dirtytalk}

\def\figdir{../Figs}

\usepackage[most]{tcolorbox}
\newcommand{\sada}[1]{{\leavevmode\color{orange}[SK: #1]}}
\newcommand{\sadaadded}[1]{{\leavevmode\color{orange}{#1}}}
\newcommand{\todo}[1]{{\leavevmode\color{orange}[TODO: #1]}}
\newcommand{\add}[1]{{\leavevmode\color{blue}#1}}
\newcommand{\del}[1]{{\leavevmode\color{red}\sout{#1}}}
\newcommand{\response}[1]{{\leavevmode\noindent #1}}
%\newcommand{\rcomment}[1]{{\bigskip\begin{snugshade}\begin{quotation}\noindent #1 \end{quotation}\end{snugshade} }}
\newcommand{\rcomment}[1]{%
\vspace{10pt}
\begin{tcolorbox}[colback=black!3,colframe=white!45!black]
#1
\end{tcolorbox}
}
\newcommand{\correction}[2]{{\begin{quotation}\noindent {#1}{\it #2} \end{quotation} }}
%\newcommand{\correction}[2]{{\bigskip\begin{quotation}\noindent {#1}{\it #2} \end{quotation} }}

\newcommand{\mycaption}[2]{%
  \caption[#1]{\textbf{#1} \small#2}%
}
% comment out when submitting
\definecolor{myblue}{RGB}{33,150,243}
\definecolor{mygreen}{RGB}{76, 175, 80}
\definecolor{purple}{RGB}{170, 0, 255}

@@

\makeatletter
\renewcommand{\maketitle}{\bgroup\setlength{\parindent}{0pt}
\begin{flushleft}
\Large  \textbf{\@title}
\end{flushleft}\egroup
}
\makeatother
\title{Response to the review of submission\\  ``Unsupervised embedding of trajectories captures the latent structure of mobility''}
\date{}
\begin{document}
\maketitle

\noindent Dear Dr.editorname and Refrees,

We would like to thank all Referees for their incredibly helpful and insight comments which have helped to improve the quality of our manuscript.
In the following, we provide detailed responses to the issues raised by each Referee.
A version of our manuscript with changes annotated has also been provided to highlight changes made.

\tableofcontents
\listoffigures

\section{List of changes}

We have broken down the list of changes to the manuscript by referee.

\section{Point-by-point discussion}

\subsection{Response to Referee 1}

@@

}

\response{
We believe that the mathematical breakthroughs made since our first submission improve the manuscript and extend its contribution and demonstrates novel scientific insight, and justifies the use of word2vec towards mobility data. By adding this, and addressing your other points (above), we believe that the revised manuscript sufficiently remedies your concerns.

}

\bigskip
\response{%
In summary, we have updated the manuscript by providing the appropriate context in the literature, clarifying the limitations of our study, and extending our approach to incorporate delays in contact tracing. We deeply appreciate the Referee's careful reading of the paper and their constructive comments. 
%We hope that have correctly addressed the Referee's concerns and comments.
}

@@

\newpage

\bigskip
\subsection{Response to Referee 2}

\rcomment{%
I have carefully read the manuscript by Murray and others entitled as “Unsupervised embedding of trajectories captures the latent structure of mobility”, submitted for publication in Nature. In this paper the authors modify a conventional embedding method originally developed for word embedding, the Word2Vec algorithm, to obtain low dimensional representation of individual mobility trajectories in different settings. They analysed an air transportation dataset describing 87)flight connections between domestic airport of the US, a South Korean hotel accommodation dataset, and a world wide scientific mobility dataset obtained from the affiliations of researchers, recorded in the ISI Web of Science corpus. Beyond the methodological advance the main result of the paper is about the latent geographical and hierarchical patterns of academic units, what the embedding built from short scientific trajectories is demonstrated to capture at once.paradox of friendship) and propose a viable way to exploit the first principle without knowing the directionality of the infections.
}

\response{%
We would like to thank the Referee for helpful comments. 
In the following, we provide detailed responses to all issues raised by the Referee.
}

@@

\response{%
Thank you for pointing this out. 
We carefully read out the Barthélemy, Physics Reports 499, 1, 2011, and referenced it in our manuscript. 
Also, we report the performance of the flux estimation (Fig.~\ref{fig:supp:distancemetrics}, this document) and prediction (Fig.~\ref{fig:supp:distancermse}, this document) using different deterrence functions, funding that while performance differs, that the embedding distance outperforms geographic 
distance no matter the function used. 
We add the following line to the methods,

\medskip
\say{The decay function $f(r_{ij})$ defines the effect of distance, and so different decay functions can model fundamentally different mechanism [...] such as the cost functions for a given distance and the spatial granularity of the observation.
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case. though the embedding distance outperforms geographic distance regardless of the type of gravity model used (see Supporting Information).}
\medskip

We also argue that geographical distance ﬁts well with the power-law decay function because it is is more suitable for analyzing large, complex, and scale-free spatial systems , while embedding distance ﬁts well with the exponential decay function, which we demonstrate to correspond to an underlying mathematical relationship between \textit{word2vec} and the gravity model, a connection we recently discovered and included as a major finding in this revision (see our response to Referee 1).

@@

}

\subsection{Response to Referee 3}

\rcomment{%
In this article manuscript, Murray et al demonstrate the power of “vanilla” embedding models based on undirected trajectories and co-affiliation to capture subtle, “functional” distances between establishments, airports, scientific universities and research institutes in terms of human mobility between them. Specifically, they embed these various institutions in a 300 dimensional Euclidean space with a shallow neural network (the very popular word2vec skip-gram architecture) that enables recovery of many subtle aspects of these spaces. For example, when the authors replace geographic distance in the well-known gravity model with cosine distance from their embedding model, they more than double their prediction of human flux between the institutions in each case as well as overall. Moreover, when they interrogate the case of scientific institutions, they show (1) through dimension-reduction (with the UMAP algorithm) on their 300D space, they can recover a high-resolution 2D map of science that captures many intrinsic dimensions of the scientific landscape (e.g., like institution types, because they are built from the mobility of people who are selected and selecting between the institutions), (2) through anchoring the dimensionality of the Euclidean space (e.g., high minus low status institutions) they can identify geographic location and status; (3) through hierarchical clustering and a measure of hierarchical “depth”, the similarity of national scientific enterprises, suggesting the multi-scale resolution of these representations. In short, they demonstrate the power of popular, easy-to-run, shallow embedding approaches to capture and represent structures related to human mobility, largely deploying interesting bespoke methods they have previously developed (and published) to navigate these representations (e.g., SemAxis and Element-Centric clustering).
}

\response{%
Thank you for the valuable comments. 
In the following, we go through the review point-by-point and address the Referee's concerns one by one.
}

@@

\rcomment{
Finally, one cannot escape the fact that COVID changes our mobility patterns. Many groups have used mobile phone data to explore these changes. I wonder, how would the pre-covid embeddings apply to post-covid mobility? Once again, I am mindful that this would require access to such data, but I would not be surprised at all if the authors have access to such data—in which case this is remains an intriguing question.
}

\response{
This is an excellent idea, and one that we have been interesting in addressing as well. 
Unfortunately, however, we do not have access to such data, at least not in large enough quantiaties to be representative. 
Publications are on a time lag, compared to actual events.
For example, a researcher may write a paper while studying at university $A$, but then before the final paper is published, move to university $B$, so while they now have a $B$ affiliation, in our data all we see is the $A$ one. 
As such, we would not necessarily expect to  see the impact of COVID on bibliometric data until later in 2021 or early 2022, when researchers have had the chance to publich with their new institution.
We also not that some activity will, of course, be visible, especially given the number of researchers rapidly publishing on the subject of COVID-19, however we consider this to be an unrepresentative sample, given that it is concentrated in a few disciplines.

In the future, however, we hope to have access to a fully-representative sample of mobility data to study the effects of COVID-19 on global scientific labor flows. 
We can use techniqeus developed by other researchers, such as those used by Hamilton, Leskovec, \& Jurafsky (\href{https://nlp.stanford.edu/projects/histwords/}{[link]}), to align embeddings trained on different data in order to quantify the changes in global mobility. 
We look forward to approaching this, as well as other exciting projects in the future. 
}

\section{Theoretical justification of \textit{word2vec} and the Gravity Model} 
\todo {add text}

\section{Performance of \textit{word2vec} compared with baselines}
\todo{add text}

\clearpage
%
% Figure - Distance metric performance
%
\begin{figure}[p!]
	\centering
	\includegraphics[width=0.9\textwidth]{\figdir/Descriptive/distance_metric_performance.pdf}
		\mycaption
		{Cosine distance between embedding vectors explains more flux in scientific mobility than geographic distance, baselines.}
		{The correlation, or amount of flux explained by the embedding distance across distance metrics and other experimental parameters. 
		All correlations are significant at $p < 0.0001$). 
		Distance metrics are ordered, from top to bottom, by the highest correlation ($r^{2}$). 
		The asterisk denotes the top-performing distance metric for every case. 
		All distance metrics explain more of the flux than does geographic distance.
		The cosine distance between embedding vectors explains more flux than all other distance metrics. 
		The x-axis differentiates between three strategies for defining the organization population used to calculate the flux. 
		``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
		population is defined in the same way for ``mobile-only'', except only using unique mobile researchers;
		``Raw frequency'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding. 
		Generally, ``All'' explains more of the flux, however there is little difference between definitions of mobility.
		All distance metrics tend to explain more of the flux when describing all or domestic (same-country) mobility.	
		}
	\label{fig:supp:distancemetrics}
\end{figure}

\clearpage
\begin{figure}[p!]
	\centering
	\includegraphics[width=0.85\textwidth]{\figdir/Descriptive/distance_prediction_performance_rmse.pdf}
		\mycaption
		{Using cosine distance between embedding vectors as input to the gravity model results in better predictions of scientific mobility compared to geographic distance, baselines.}
		{Shown is the Root Mean Squared Error (RMSE), or the accuracy of predictions made with the gravity model of mobility using the embedding distance and other experimental parameters.
		The RMSE is shown using both the exponential and power-law forms of the gravity model. 
		For each distance metric and set of parameters, the RMSE is shown for 
		Distance metrics are ordered, from top to bottom, by the lowest RMSE, where a lower error indicates better predictions. 
		The asterisk denotes the best-performing distance metric for every case. 
		The x-axis differentiates between three strategies for defining the organization population used to calculate the flux. 
		``All'' means that population is defined as the average yearly number of unique mobile and non-mobile scholars who published with the organizations' affiliation;
		population is defined in the same way for ``mobile-only'', except only using unique mobile researchers;
		``Raw frequency'' means that organization populations are defined as their frequency across all the trajectories, similar to word frequency in language embedding. 
		Generally, ``All'' results in better prediction.
		}
	\label{fig:supp:distancermse}
\end{figure}

### 2021-02-04 paper/rebuttal/rebuttal.tex

\documentclass[12pt,a4paper]{article}
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage[margin=1in]{geometry}
\usepackage{enumitem}
\usepackage{pslatex}
\usepackage{color}
\usepackage{bm}
\usepackage{xcolor}
\usepackage[normalem]{ulem}
\definecolor{orange}{RGB}{230, 81, 0}
\definecolor{purple}{RGB}{170, 0, 255}
\definecolor{mygreen}{RGB}{76, 175, 80}
\usepackage{framed}
\definecolor{shadecolor}{RGB}{210, 210, 210}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{newtxmath}
\DeclareMathAlphabet{\mathpzc}{T1}{pzc}{m}{it}
\def\given{\mid}
\usepackage{bm}
\def\tnull{{\text{null}}}
\def\vec#1{{\bm #1}}
\def\mat#1{\mathbf{#1}}

% abbreviations
\def\etal{\emph{et~al}.\ }
\def\eg{e.g.,~} 
\def\ie{i.e.,~}
\def\cf{cf.\ }
\def\viz{viz.\ }
\def\vs{vs.\ }

\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=black,
    filecolor=magenta,      
    urlcolor=cyan,
}
\usepackage[style=nature,
					backend=biber,
					sortcites=true,
					autocite=superscript
]{biblatex}
\addbibresource{rebuttal.bib}

\DeclareUnicodeCharacter{FB01}{fi}
\usepackage{dirtytalk}

@@

\response{%
Thank you for the valuable comments. 
In the following, we go through the review point-by-point and address the Referee's concerns one by one.
}

\rcomment{
The analysis does not compare embeddings, which would not be a problem if they were making
an inference other than we can embed micro-mobility to trace macro-mobility. They use an
efficient (Google engineers developed it), single-layer embedding from 2013, which produces
linear associations. But the paper doesn’t take advantage of the revolution in deep neural
network models that now produce powerful, nonlinear and noncontiguous (e.g., using the
attention architecture) models that now dominate analysis of language, and are making very
power inroads in machine vision. To be clear, deeper models require much data and computational power to tune—they are an engineering feat—but if developed in these settings, they would likely dramatically improve the prediction of flux. They take into account directionality (e.g., a person moving from one opportunity to another) and would independently capture people how people distinctly move up and down the hierarchy or from place to place (e.g., a in bi-directional model like BERT). A lack of discussion or demonstration of which model works best--and the unfounded inference that direction does not matter felt weak without a demonstration that they would not have worked better. They did and do work much better for language models and for image generation...why would they not work for analysis of human mobility? (I’m sure they would, but this would time and independent development.)

}

\response{
Thank you for this insightful comment. 
We think that there is a great deal of potential in methods like BERT for representing mobility data.
What initially drew us to \textit{word2vec}, however, was its simplicity, its ubiquity, and its mathematical tractability (at least compared to other, more sophisticated approaches). 
This last attribute proved very useful, as in exploring the mathematical properties of the model, we discovered a deep connection between \textit{word2vec} and the Gravity Model of Mobility, which explains why it learns such effective gravity-like relationships between entities. 
We believe hat this new discovery, not included in the manuscript (and illustrated below), justify our use of \textit{word2vec} here. 
We agree that the revolution in neural embeddings offers great promise in mobility and look forward to reading and also working on these future studies; however we argue that our contribution is most well founded, and best serves the scientific community, by working with simpler and more established models, especially concerning the mathematical link we uncovered. 
To highlight the importance of altenrative models, however, we add the following line to our conclusion: \say{\textit{Future studies may also consider bi-directional embeddings, such as BERT~\autocite{devlin2018bert}, to incorporate directionality. }}.

@@

\rcomment{
Finally, one cannot escape the fact that COVID changes our mobility patterns. Many groups have used mobile phone data to explore these changes. I wonder, how would the pre-covid embeddings apply to post-covid mobility? Once again, I am mindful that this would require access to such data, but I would not be surprised at all if the authors have access to such data—in which case this is remains an intriguing question.
}

\response{
This is an excellent idea, and one that we have been interesting in addressing as well. 
Unfortunately, however, we do not have access to such data, at least not in large enough quantiaties to be representative. 
Publications are on a time lag, compared to actual events.
For example, a researcher may write a paper while studying at university $A$, but then before the final paper is published, move to university $B$, so while they now have a $B$ affiliation, in our data all we see is the $A$ one. 
As such, we would not necessarily expect to  see the impact of COVID on bibliometric data until later in 2021 or early 2022, when researchers have had the chance to publich with their new institution.
We also not that some activity will, of course, be visible, especially given the number of researchers rapidly publishing on the subject of COVID-19, however we consider this to be an unrepresentative sample, given that it is concentrated in a few disciplines.

In the future, however, we hope to have access to a fully-representative sample of mobility data to study the effects of COVID-19 on global scientific labor flows. 
We can use techniqeus developed by other researchers, such as those used by Hamilton, Leskovec, \& Jurafsky (\href{https://nlp.stanford.edu/projects/histwords/}{[link]}), to align embeddings trained on different data in order to quantify the changes in global mobility. 
We look forward to approaching this, as well as other exciting projects in the future. 
}

\section{Theoretical justification of \textit{word2vec} and the Gravity Model} 
In here, we would like to answer a question: why we use SGNS algorithms and is there any theoretical justification on that? This document details:
\begin{enumerate}
	\item Details on \textit{word2vec} algorithm

\end{enumerate}

\subsection{ Details on \textit{word2vec} algorithm}
We used standard skip-gram negative sampling word embedding, commonly known as \textit{word2vec}~\autocite{mikolov2013word2vec}.
Suppose a trajectory, denoted by ($a_{1}, a_{2}, \ldots, a_{T}$), where $a_{t}$ is the $t$th location in the trajectory. A location, $a_{t}$, is considered to have context locations, $a_{t-w}, \ldots, a_{t-1}, a_{t+1},\ldots, a_{t+w}$, that appear in the window surrounding $a_t$ up to a time lag of $w$, where $w$ is the window size parameter truncated at $t - w \geq 0$ and $t + w \leq T$. Then, the model learns probability $p(a_{t + \tau} \vert a_{t})$, where $-w\leq \tau\leq w$ and $\tau \neq 0$,  by maximizing its log likelihood given by
%
% MAIN W2V Equation
%
\begin{equation}
	\frac{1}{T}\sum_{t = 1}^{T} \sum_{-w \leq \tau \leq w, \tau \neq 0} \log p(a_{t + \tau} \vert a_{t}),
\end{equation}
where,

%
% CONDITIONAL PROBABILITY EQ
%
\begin{equation}
	p(j \given i) = \frac{\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i}, \label{eq:cond_prob_w2v}
\end{equation}
where $\bm{v}$ and $\bm{u}$ are the ``in-vector" and ``out-vector", respectively,  $Z_i=\sum_{j' \in \mathcal{A}} \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$ is a normalization constant, and $\mathcal{A}$ is the set of all locations. In general, calculating $Z_i$ is computationally expensive and there are two common approximation: 
 hierarchical softmax \autocite{morin2005hierarchical} and negative sampling \autocite{mikolov2013word2vec}.

\subsection{An implicit bias in the negative sampling}

In our paper, we employed a specific algorithm SGNS word embedding. However, it is well known that the negative sampling is a biased estimator~\autocite{Chia2010,Dyer2014} that optimizes a distinct model  from {\it word2vec}.

Negative sampling trains {\it word2vec} using a binary classification task as follows.
A center-context word pair $(i,j)$ is sampled from the given data as a positive pair, and $k$ negative pairs $(i,\ell)$ are sampled from a noise distribution $p_0(\ell)=P^\gamma(\ell)$. ($P(j)$ is the fraction of $j$ in the data, and $\gamma>0$ is a parameter).
Then, given a center-context word pair $(i,j)$, the model predicts whether it is a positive or negative sample using sigmoid function:
\begin{align}
	\label{eq:sigmoid}
	\frac{1}{1 + \exp\left( - \bm{u}_j \cdot \bm{v}_i \right)}.
\end{align}

This procedure does not guarantee that the embedding optimally converges to the original conditional probability eq.\ref{eq:cond_prob_w2v}, even when increasing the training samples and iterations~\autocite{Chia2010,Dyer2014} in the fact that negative sampling has implicit bias.
This problem is addressed by the noise contrastive estimation (NCE) \autocite{Chia2010}, which fits a probability model in the form:
\begin{align}
	\label{eq:nce}
	f(x) / \sum_{x'} f(x'),
\end{align}
where $f(x)$ is the unnormalized probability for an instance $x$. As in negative sampling, NCE fits the model based on a binary classification task, with positive and negative samples randomly drawn from the given data and a noise distribution $p_0$, respectively, using a bias-corrected sigmoid function:
\begin{align}
	\label{eq:sigmoid2}
	\frac{1}{1 + \exp\left( -f(x) + \log p_0(x) \right)}.
\end{align}

Rewriting the sigmoid function for the negative sampling (Eq.~\eqref{eq:sigmoid}) in form of the bias-corrected sigmoid function yields
\begin{align}
	\label{eq:sigmoid3}
	\frac{1}{1 + \exp\left[ - \tilde f(x) + \log p_0(x)  \right]},
\end{align}
where we have defined
\begin{align}
	\label{eq:unnormalized}
	\tilde f(x) = f(x) p_0(x).
\end{align}
So, the model with the negative sampling optimize $\tilde f(x)$, not $f(x)$. In case of the {\it word2vec}, the instance is the dot similarity $x=\bm{u}_i \cdot \bm{v}_j$,  $f(x)=\exp(x)$, and $p_0(j)=P^\gamma(j)$, and the SGNS word2vec actually optimize the following conditional probability:
\begin{align}
	P\left(j \given i \right):= \frac{P^\gamma (j)\exp(\bm{u}_j \cdot \bm{v}_{i})}{\widetilde{Z_i}},
\end{align}
where $\widetilde {Z_i}=\sum_{j' \in \mathcal{A}} P^\gamma (j') \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$.

\subsection*{word2vec and the gravity model}

Why does the embedding distance predict actual mobility flows better than the geographic distance?
Here, we show that the objective function of {\it word2vec} is equivalent to the gravity model.
 {\it Word2vec} learns an embedding by estimating the probability that location $i$ has context $j$:
\begin{align}
	P\left(j \given i \right):= \frac{\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i}, \label{eq:prob_w2v_ng}
\end{align}
where the denominator $Z_i=\sum_{j' \in \mathcal{A}} \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$ is a normalization constant, and $\mathcal{A}$ is the set of all locations.
Although {\it word2vec} generates two embedding vectors $\bm{v_i}$ and $\bm{u}_i$---which are referred to as the in-vector and out-vector, respectively, we follow many studies by adopting the in-vector $\bm{v}_i$ as an embedding of location $i$.

Our experiments show that {\it word2vec} best explains real-world mobility flow when $w=1$ (Fig.~\ref{fig:supp:hyperparameter}), with the flow predicted by
\begin{align}
	\label{eq:flow_w2v}
	\hat{T_{ij}}= P(i) P(j\given i) = \frac{P(i)\exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i},
\end{align}
where $P(i)$ is the probability that location $i$ appears in the trajectory.
Because computing $Z_i$ is computationally demanding, multiple approximation methods have been proposed, the most  popular being negative sampling~\autocite{mikolov2013word2vec}. 
Negative sampling finds the embedding vectors by solving a binary classification task, in which the model classifies whether a given center-context location pair $(i, j)$ comes from the data or artificially generated ``negative'' samples.
The negative pair $(i,j)$ is sampled in proportion to $P^\gamma(j)$ with parameter $\gamma \geq 0$.

As a heuristic, the negative sampling has a bias and maximizes a different objective function~\autocite{Gutmann2010}.
When taking into account this bias, {\it word2vec} with the negative sampling fits a probability model given by
\begin{align}
	P\left(j \given i \right):= \frac{P^ \gamma(j) \exp(\bm{u}_j \cdot \bm{v}_{i})}{Z_i}, \label{eq:prob_w2v_ng0}
\end{align}
where we redefine the normalization constant as $Z_i=\sum_{j' \in \mathcal{A}} P^ \gamma (j') \exp(\bm{u}_{j'} \cdot \bm{v}_{i})$. (See Methods section for the full calculation).

Parameter $\gamma=1$ is a special choice that ensures that, when the embedding dimension is sufficiently large, there exists optimial in-vectors and out-vectors such that $\bm{u}_i = \bm{v}_i$ \autocite{levy2014neural}.
Setting $\gamma = 1$ and substituting $\bm{u}_i = \bm{v}_j$ into Eq.~\eqref{eq:prob_w2v_ng0}, the flow predicted by {\it word2vec} is given by
\begin{align}
	\hat{T_{ij}}= P(i) P(j\given i) = \frac{P(i)P(j)\exp(\bm{v}_j \cdot \bm{v}_{i})}{Z_i}. \label{eq:flow_w2v_ng1}
\end{align}
The flow $\hat{T_{ij}}$ is symmetric (\ie $\hat{T_{ij}}=\hat{T_{ji}}$) because {\it word2vec} neglects whether the context $j$ appears before or after the center $i$ in the trajectory.
If we swap $i$ and $j$ in Eq.~\eqref{eq:flow_w2v_ng1}, the numerator remains the same but the denominator can be different.
Therefore, to ensure $T_{ij} = T_{ji}$, the denominator $Z_i$ should be a constant.

Taken altogether, the {\it word2vec} model with the negative sampling predicts a flow in the same form as in the gravity model:
\begin{align}
	\hat{T_{ij}}=  C P(i) P(j)  \exp(\bm{v}_j \cdot \bm{v}_{i}), \label{eq:flow_w2v_ng0}
\end{align}
where $C$ is a positive constant.
In other words, \emph{word2vec with the negative sampling is a gravity model}, with the mass given by the location's frequency $P(i)$, and the distance measured by their dot similarities.
While the gravity model generates the flow from the given mass and locations, {\it word2vec} finds the mass and locations that best explain the given trajectories.
Therefore, the embedding distance is inherently tied to the flow, and hence, has greater predictive power than the geographic distance.

### 2021-02-10 paper/current/main.tex
add text about why cosine better and in/out are different

>>>>

A central assumption in our mathematical analysis is that the embedding vectors are optimum in terms of the {\it word2vec}'s objective function.
However, the embedding vectors are sub-optimal in practice because the optimization algorithm is stochastic and can be trapped in a local optimum.
This has two consequences.
First, the suboptimal in- and out-vectors can be different, which goes against the premise that the (optimal) in-vectors and out-vectors are identical.
Second, dot similarity should in principle predict the mobility flow better than cosine similarity because dot similarity includes cosine similarity as its special case.
However, our simulations show that it is the cosine similarity that better predicted the mobility flow~[reference to a figure]. This phenomenon---cosine similarity outperforms dot similarity---has been observed in many applications of word embedding~\cite{add citations}. Yet, why cosine similarity is superior to dot similarity has not been clearly understood.
Nevertheless, among the 7(?) measures tested in our simulations, dot similarity is the second-best predictor of mobility flow following cosine similarity. Furthermore, unlike cosine similarity, dot similarity has a mathematical connection to the gravity model.
